---
title: "HCC Risk Score Time Series Analysis"
author: "Amy Jin"
date: "1/22/2019"
output: html_document
---

## Risk Score Data Processing

Data source: https://data.cms.gov/browse?tags=risk%20score

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
d2014 <- read.csv("/Users/yuejin/Desktop/HCC_Risk_Score_Analysis/2014_County-level_Fee_for_Service__FFS__Data_for_Shared_Savings_Program__SSP__Benchmark_Rebasing_PUF.csv", header =T)
d2015 <- read.csv("/Users/yuejin/Desktop/HCC_Risk_Score_Analysis/2015_County-level_Fee_for_Service__FFS__Data_for_Shared_Savings_Program__SSP__Benchmark_Rebasing_PUF.csv", header =T)
d2016 <- read.csv("/Users/yuejin/Desktop/HCC_Risk_Score_Analysis/2016_County-level_Fee_for_Service__FFS__Data_for_Shared_Savings_Program__SSP__Benchmark_Rebasing_PUF.csv", header =T)
d2017 <- read.csv("/Users/yuejin/Desktop/HCC_Risk_Score_Analysis/2017_County-level_Fee_for_Service__FFS__Data_for_Shared_Savings_Program__SSP__Benchmark_Rebasing_PUF.csv", header =T)
```

```{r}
#colnames(d2014)
library(plyr)
d1415 <- join(d2014, d2015, by=c("State.ID", "County.ID"), type="inner")
d141516 <- join(d1415, d2016, by=c("State.ID", "County.ID"), type="inner")
d14151617 <- join(d141516, d2017, by=c("State.ID", "County.ID"), type="inner")
#colnames(d14151617)
join_table <- d14151617[,c(2,3,4,5,1,7,18,22,33,37,48,52)]
#head(join_table)
# There are . * in Average.ESRD.HCC.risk.score.3
join_table$Average.ESRD.HCC.risk.score.3[join_table$Average.ESRD.HCC.risk.score.3 == '.'] <- NA
join_table$Average.ESRD.HCC.risk.score.3[join_table$Average.ESRD.HCC.risk.score.3 == '*'] <- NA

# Difference between two years
join_table$Average.ESRD.HCC.risk.score.3 <- as.numeric(as.character(join_table$Average.ESRD.HCC.risk.score.3))
join_table$diff15_14 <- join_table$Average.ESRD.HCC.risk.score.1 - join_table$Average.ESRD.HCC.risk.score
join_table$diff16_15 <- join_table$Average.ESRD.HCC.risk.score.2 - join_table$Average.ESRD.HCC.risk.score.1
join_table$diff17_16 <- join_table$Average.ESRD.HCC.risk.score.3 - join_table$Average.ESRD.HCC.risk.score.2

join_table$diff16_14 <- join_table$Average.ESRD.HCC.risk.score.2 - join_table$Average.ESRD.HCC.risk.score
join_table$diff17_15 <- join_table$Average.ESRD.HCC.risk.score.3 - join_table$Average.ESRD.HCC.risk.score.1

# Rate between two years
join_table$pct15_14 <- join_table$diff15_14/join_table$Average.ESRD.HCC.risk.score
join_table$pct16_15 <- join_table$diff16_15/join_table$Average.ESRD.HCC.risk.score.1
join_table$pct17_16 <- join_table$diff17_16/join_table$Average.ESRD.HCC.risk.score.2
head(join_table)
```

## Hitograms: Average ESRD HCC Risk Score Difference

```{r, fig.width=10}
hist(join_table$diff15_14, 
     main="Histogram: Average ESRD HCC Risk Score Difference between 2015 and 2014", 
     xlab="Average ESRD HCC Risk Score Difference", 
     border="blue", 
     col="green",
     xlim=c(-1,1),
     las=1, 
     breaks=50)
```


```{r, fig.width=10}
hist(join_table$diff16_15, 
     main="Histogram: Average ESRD HCC Risk Score Difference between 2016 and 2015", 
     xlab="Average ESRD HCC Risk Score Difference", 
     border="blue", 
     col="green",
     xlim=c(-1,1),
     las=1, 
     breaks=50)
```


```{r, fig.width=10}
hist(join_table$diff17_16, 
     main="Histogram: Average ESRD HCC Risk Score Difference between 2017 and 2016", 
     xlab="Average ESRD HCC Risk Score Difference", 
     border="blue", 
     col="green",
     xlim=c(-1,1),
     las=1, 
     breaks=50)
```

```{r, fig.width=10}
hist(join_table$diff16_14, 
     main="Histogram: Average ESRD HCC Risk Score Difference between 2016 and 2014", 
     xlab="Average ESRD HCC Risk Score Difference", 
     border="blue", 
     col="green",
     xlim=c(-1,1),
     las=1, 
     breaks=50)
```

```{r, fig.width=10}
hist(join_table$diff17_15, 
     main="Histogram: Average ESRD HCC Risk Score Difference between 2017 and 2015", 
     xlab="Average ESRD HCC Risk Score Difference", 
     border="blue", 
     col="green",
     xlim=c(-1,1),
     las=1, 
     breaks=50)
```

## Hitograms: Average ESRD HCC Risk Score Change Rate

```{r, fig.width=10}
hist(join_table$pct15_14, 
     main="Histogram: Average ESRD HCC Risk Score Change Rate between 2015 and 2014", 
     xlab="Average ESRD HCC Risk Score Change Rate", 
     border="blue", 
     col="green",
     xlim=c(-1,1),
     las=1, 
     breaks=50)
```

```{r, fig.width=10}
hist(join_table$pct16_15, 
     main="Histogram: Average ESRD HCC Risk Score Change Rate between 2016 and 2015", 
     xlab="Average ESRD HCC Risk Score Change Rate", 
     border="blue", 
     col="green",
     xlim=c(-1,1),
     las=1, 
     breaks=50)
```

```{r, fig.width=10}
hist(join_table$pct17_16, 
     main="Histogram: Average ESRD HCC Risk Score Change Rate between 2017 and 2016", 
     xlab="Average ESRD HCC Risk Score Change Rate", 
     border="blue", 
     col="green",
     xlim=c(-1,1),
     las=1, 
     breaks=50)
```

```{r}
mean(abs(join_table$diff15_14), na.rm=TRUE)
mean(abs(join_table$diff16_15), na.rm=TRUE)
mean(abs(join_table$diff17_16), na.rm=TRUE)
```

* Mean of absolute Average ESRD HCC Risk Score difference from 2014 to 2015 is 0.04141753. 
* Mean of absolute Average ESRD HCC Risk Score difference from 2015 to 2016 is 0.04096563. 
* Mean of absolute Average ESRD HCC Risk Score difference from 2016 to 2017 is 0.04233740. 

```{r}
mean(abs(join_table$pct15_14), na.rm=TRUE)
mean(abs(join_table$pct16_15), na.rm=TRUE)
mean(abs(join_table$pct17_16), na.rm=TRUE)
```

* Mean of absolute Average ESRD HCC Risk Score change rate from 2014 to 2015 is 4.171 %. 
* Mean of absolute Average ESRD HCC Risk Score change rate from 2015 to 2016 is 4.140 %. 
* Mean of absolute Average ESRD HCC Risk Score change rate from 2016 to 2017 is 4.284 %. 

Conclusion: there is a moderate difference of Average ESRD HCC Risk Score from year to year.