---
title: "NPI Level HCC Risk Score Time Series Analysis"
author: "Amy Jin"
date: "1/23/2019"
output: html_document
---

## Risk Score Data Processing

Data Source:  https://data.cms.gov/Medicare-Physician-Supplier/Medicare-Physician-and-Other-Supplier-National-Pro/85jw-maq9

```{SQL, eval = F, results=F}
SELECT
  d13.npi                                     AS npi,
  d13.average_hcc_risk_score_of_beneficiaries AS risk_2013,
  d14.average_hcc_risk_score_of_beneficiaries AS risk_2014,
  d15.average_hcc_risk_score_of_beneficiaries AS risk_2015,
  d16.beneficiary_average_risk_score          AS risk_2016,
  d14.average_hcc_risk_score_of_beneficiaries - d13.average_hcc_risk_score_of_beneficiaries AS diff_2014_2013,
  d15.average_hcc_risk_score_of_beneficiaries - d14.average_hcc_risk_score_of_beneficiaries AS diff_2015_2014,
  d16.beneficiary_average_risk_score - d15.average_hcc_risk_score_of_beneficiaries AS diff_2016_2015,

  d15.average_hcc_risk_score_of_beneficiaries - d13.average_hcc_risk_score_of_beneficiaries AS diff_2015_2013,
  d16.beneficiary_average_risk_score - d14.average_hcc_risk_score_of_beneficiaries AS diff_2016_2014,

  (d14.average_hcc_risk_score_of_beneficiaries - d13.average_hcc_risk_score_of_beneficiaries)/d13.average_hcc_risk_score_of_beneficiaries AS pct_2014_2013,
  (d15.average_hcc_risk_score_of_beneficiaries - d14.average_hcc_risk_score_of_beneficiaries)/d14.average_hcc_risk_score_of_beneficiaries AS pct_2015_2014,
  (d15.average_hcc_risk_score_of_beneficiaries - d13.average_hcc_risk_score_of_beneficiaries)/d13.average_hcc_risk_score_of_beneficiaries AS pct_2015_2013,

  (d15.average_hcc_risk_score_of_beneficiaries - d13.average_hcc_risk_score_of_beneficiaries)/d13.average_hcc_risk_score_of_beneficiaries AS pct_2015_2013,
  (d16.beneficiary_average_risk_score - d14.average_hcc_risk_score_of_beneficiaries)/d14.average_hcc_risk_score_of_beneficiaries AS pct_2016_2014,

  (d16.beneficiary_average_risk_score - d14.average_hcc_risk_score_of_beneficiaries)/d14.average_hcc_risk_score_of_beneficiaries AS pct_2016_2014,

  (d16.beneficiary_average_risk_score - d15.average_hcc_risk_score_of_beneficiaries)/d15.average_hcc_risk_score_of_beneficiaries AS pct_2016_2015
FROM npi_summary.npi_summary_2013 AS d13
  JOIN npi_summary.npi_summary_2014 AS d14
    ON d13.npi = d14.npi
  JOIN npi_summary.npi_summary_2015 AS d15
    ON d13.npi = d15.npi
  JOIN npi_summary.npi_summary_2016 AS d16
    ON d13.npi = d16.npi;
```


```{r}
data <- read.csv("/Users/yuejin/Desktop/HCC_Risk_Score_Analysis/Risk_Score_NPI_2013_2016.csv", header = T)

head(data)
```


## Hitograms: ESRD HCC Risk Score Change

```{r, fig.width=10}
hist(data$diff_2014_2013, 
     main="Histogram: ESRD HCC Risk Score Change between 2014 and 2013", 
     xlab="ESRD HCC Risk Score Change", 
     border="blue", 
     col="green",
     xlim=c(-1,1),
     las=1, 
     breaks=500)
```


```{r, fig.width=10}
hist(data$diff_2015_2014, 
     main="Histogram: ESRD HCC Risk Score Change between 2015 and 2014", 
     xlab="ESRD HCC Risk Score Change", 
     border="blue", 
     col="green",
     xlim=c(-1,1),
     las=1, 
     breaks=500)
```


```{r, fig.width=10}
hist(data$diff_2016_2015, 
     main="Histogram: ESRD HCC Risk Score Change between 2016 and 2015", 
     xlab="ESRD HCC Risk Score Change", 
     border="blue", 
     col="green",
     xlim=c(-1,1),
     las=1, 
     breaks=500)
```

```{r, fig.width=10}
hist(data$diff_2015_2013, 
     main="Histogram: ESRD HCC Risk Score Change between 2015 and 2013", 
     xlab="ESRD HCC Risk Score Change", 
     border="blue", 
     col="green",
     xlim=c(-1,1),
     las=1, 
     breaks=500)
```

```{r, fig.width=10}
hist(data$diff_2016_2014, 
     main="Histogram: ESRD HCC Risk Score Change between 2016 and 2014", 
     xlab="ESRD HCC Risk Score Change", 
     border="blue", 
     col="green",
     xlim=c(-1,1),
     las=1, 
     breaks=500)
```

## Hitograms: ESRD HCC Risk Score Percentage Change

```{r, fig.width=10}
hist(data$pct_2014_2013, 
     main="Histogram: ESRD HCC Risk Score Percentage Change from 2013 to 2014", 
     xlab="ESRD HCC Risk Score Change Rate", 
     border="blue", 
     col="green",
     xlim=c(-1,1),
     las=1, 
     breaks=500)
``` 

```{r, fig.width=10}
hist(data$pct_2015_2014, 
     main="Histogram: ESRD HCC Risk Score Percentage Change from 2014 to 2015", 
     xlab="ESRD HCC Risk Score Percentage Change", 
     border="blue", 
     col="green",
     xlim=c(-1,1),
     las=1, 
     breaks=500)
```

```{r, fig.width=10}
hist(data$pct_2016_2015, 
     main="Histogram: ESRD HCC Risk Score Percentage Change from 2015 to 2016", 
     xlab="HCC Risk Score Percentage Change", 
     border="blue", 
     col="green",
     xlim=c(-1,1),
     las=1, 
     breaks=500)
```

```{r, fig.width=10}
hist(data$pct_2015_2013, 
     main="Histogram: ESRD HCC Risk Score Percentage Change from 2013 to 2015", 
     xlab="ESRD HCC Risk Score Percentage Change", 
     border="blue", 
     col="green",
     xlim=c(-1,1),
     las=1, 
     breaks=500)
```

```{r, fig.width=10}
hist(data$pct_2016_2014, 
     main="Histogram: ESRD HCC Risk Score Percentage Change from 2014 to 2016", 
     xlab="HCC Risk Score Percentage Change", 
     border="blue", 
     col="green",
     xlim=c(-1,1),
     las=1, 
     breaks=500)
```

```{r}
mean(abs(data$diff_2014_2013), na.rm=TRUE)
mean(abs(data$diff_2015_2014), na.rm=TRUE)
mean(abs(data$diff_2016_2015), na.rm=TRUE)
mean(abs(data$diff_2015_2013), na.rm=TRUE)
mean(abs(data$diff_2016_2014), na.rm=TRUE)
```

* Mean of absolute ESRD HCC Risk Score change from 2014 to 2015 is 0.1575151. 
* Mean of absolute ESRD HCC Risk Score change from 2015 to 2016 is 0.1719135. 
* Mean of absolute ESRD HCC Risk Score change from 2016 to 2017 is 0.1621504. 
* Mean of absolute ESRD HCC Risk Score change from 2015 to 2013 is 0.1972525. 
* Mean of absolute ESRD HCC Risk Score change from 2016 to 2014 is 0.208877. 

```{r}
mean(abs(data$pct_2014_2013), na.rm=TRUE)
mean(abs(data$pct_2015_2014), na.rm=TRUE)
mean(abs(data$pct_2016_2015), na.rm=TRUE)
mean(abs(data$pct_2015_2013), na.rm=TRUE)
mean(abs(data$pct_2016_2014), na.rm=TRUE)
```

* Mean of absolute ESRD HCC Risk Score change percentage from 2014 to 2015 is 10.45657 %. 
* Mean of absolute ESRD HCC Risk Score change percentage from 2015 to 2016 is 11.83507 %. 
* Mean of absolute ESRD HCC Risk Score change percentage from 2016 to 2017 is 10.41507 %. 
* Mean of absolute ESRD HCC Risk Score change percentage from 2015 to 2013 is 13.27411 %. 
* Mean of absolute ESRD HCC Risk Score change percentage from 2016 to 2014 is 14.30579 %. 

```{r}
sd(abs(data$pct_2014_2013), na.rm=TRUE)
sd(abs(data$pct_2015_2014), na.rm=TRUE)
sd(abs(data$pct_2016_2015), na.rm=TRUE)
```

* Standard Deviation of absolute ESRD HCC Risk Score change percentage from 2014 to 2015 is 0.1244735. 
* Standard Deviation of absolute ESRD HCC Risk Score change percentage from 2015 to 2016 is 0.1346847. 
* Standard Deviation of absolute ESRD HCC Risk Score change percentage from 2016 to 2017 is 0.1256690. 

Conclusion: there is a moderate difference of Average ESRD HCC Risk Score from year to year.